An Expressive and Efficient Solution to the Service Selection Problem

### 2010-06-28 publicaciones/iswc10/paperISWC10v2.tex
spellchecking del paper v2

We propose a solution that extends the recent approach of Arvelo et
al.\ \cite{arvelo:aaai06} which is based on the efficient enumeration
of models of a propositional logic  theory. Similarly, given an
instance of SSP, a logical theory is constructed such  that each model
of the theory encodes a valid rewriting; the theory can be con-
structed in polynomial time from the instance of the SSP. All valid
rewritings  are obtained by enumerating all models of the theory in
linear time by using  off-the-shelf model enumeration solvers such as
c2d~\cite{c2d} and Relsat~\cite{relsat}; off-the-  seft SAT solvers
such as Minisat~\cite{minisat} can be used to produce all the minimal
rewritings. Thus, we exploit known properties of modern SAT solvers
that  implement non-chronological backtacking via clause learning,
caching of common  subproblems and heuristics, to provide an efficient
and scalable solution to SSP.

@@

Our setting is based on the assuption that services are
semantically described in terms of a domain ontology. 
Formally, the integration framework corresponds to a tuple
$IS=\tup{D,S,M}$ where $D$ is the domain ontology, $S$ describes
the available services, and $M$ is the collection of LAV mappings
that define the services in $S$ as views on the ontology $D$. 
On the other hand, the user input $R=\tup{Q,P}$ consists of a
request for composition of services, described as a conjunctive
query $Q$ over the ontology, and a set $P$ of preferences for
the requested composition.

@@

Although limited in appearance, rules of this form are quite expressive
as they allow us to specify diverse relationships among predicates.
For example,
\begin{itemize}
\item \textbf{Hierarchy of classes and subclasses} (or types and subtypes):
classes are specified with unary predicates. A subclass relantionship
among two classes is directly specified with a simple rule; e.g., the
rule $\textit{penguin}(x) \orule \textit{bird}(x)$ declares that every
penguin is a bird.
\item \textbf{Subrelations via specialization:} A subrelation of $R^r$
can be specified by constraining another relation $P^s$ ($r<s$). 
For example, the rule:
\begin{equation}
\label{eq:QE2}
\textit{descendant}(\textit{Elizabeth II},x)\ \orule\ \textit{noble}(x)
\end{equation}
tells that any descendant of Queen Elizabeth II is a noble.
\item \textbf{Indirect subsumption:} it is even possible to specify
a subrelation via another seemengly unrelated predicate. For example,
the rule:
\[ \textit{citizen-of}(x,\textit{Montreal})\ \orule\ \textit{lives-in}(x,\textit{Canada}) \]
says that when the second argument of `CitizenOf' is fixed to the
constant `Montreal', the tuples in the relation `CitizenOf' are
contained in the set of tuples in the relation `LivesIn' whose
second component is `Canada'.
\end{itemize}
However, we require that the \emph{dependency graph} for the ontology
to be a \emph{forest of trees}. The dependency graph $G(D)$ is a
labeled directed graph that is constructed as follows: the nodes
are the relational symbols in the signature $\sigma$, and there
is an edge $(R,P)$ in the graph iff there is an rule of the
form: \eqref{eq:orule} that is labeled with the \emph{bindings}
induced by the rule. For example, if $\textit{descendant}(x,y)$
and $\textit{noble}(z)$ are two predicates in the signature
and there is the rule \eqref{eq:QE2}, there edge from 
Descendant to Noble is labeled with the bindings
$\{x=\textit{Elizabeth II},y=z\}$.

@@

\Omit{
$S$ is a signature  defined as a tuple $<$$S_1^{j1}$,$\dots$,
$S_q^{jn}$, $d_1$,$\dots$,$d_k$$>$, where $S_i^{ji}$ corresponds
to a service predicate of arity $ji$ and $d_i$ is a constant.  
Mappings in $M$ correspond to semantic descriptions expressed
following the  Local as View paradigm (LAV). We assume that
services are not complete, i.e., every tuple produced by the
source in the head of the rule must satisfy the body, but not
viceversa. LAV rules are commonly used in integration systems
in order to scale up to new sources, in contrast to the Global
As View paradigm (GAV) where domain ontology predicates are
defined in terms of conjunctive formulas in the service predicates.
In addition, in LAV mappings, services are treated independently
and the integration system easily scales up to the evolution of
the available services. However, the problem of rewriting a query
on the domain ontology into a set of queries on the available
services, corresponds to the problem of rewriting queries using
views which is known to be NP-complete~\cite{RajaramanSU95,Ullman00}.  
}

@@

Let us consider a simple travel-information system that contains
information about flight and train trips between cities and information
about which cities are in the US. The domain ontology is comprised
of the predicates $\trip^2$, $\AAflight^2$, $\UAflight^2$, $\ATtrain^2$,
$\UPtrain^2$, $\flight^3$, $\train^3$ and $\UScity^1$, and the constants
$\AA$, $\UA$, $\AT$, $\UP$, $\LA$, $\NY$, $\PA$.
The first predicate relates cities $(x,y)$ if there is a direct
trip either by plane or train between them.
The predicates $\AAflight$ and $\UAflight$ relate cities that are
connected by direct flights either with American (AA) or United (UA).
The predicates $\ATtrain$ and $\UPtrain$ relate cities that are
connected by direct train trips either with Amtrak (AT) or Union
Pacific Railway (UP). The flight predicate relates $(x,y,t)$ whenever
there is a direct flight from $x$ to $y$ operated by airline $t$,
and similarly for $\train$. Finally, $\UScity$ indicates when a
given city is a US city or not.
There are various subsumption relationships among these predicates
that are captured by the following axioms:
\begin{alignat*}{1}
\flight(x,y,t)\  &\orule\ \trip(x,y)\,, \\
\train(x,y,t)\   &\orule\ \trip(x,y)\,.
\end{alignat*}
For the services, assume that the available data sources on the
Internet contain the following information:
\begin{enumerate}[--]
\item $\nationalFlight(x,y)$ relates two US cities that are connected by a direct flight,
\item $\onewayFlight(x,y)$ relates two cities that are connected by a one-way flight,
\item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
\item $\toPA(x)$ tells if there is a direct flight from $x$ to Paris,
\item $\onestopPA(x,y)$ if there is a flight from $x$ to Paris with a stop at $y$, 
\item $\fromLA(x)$ tells if there is a flight from Los Angeles to $x$,
\item $\nationalTrain(x,y)$ relates US cities that are connected by a direct train, and
\item $\onewayTrain(x,y)$ relates US cities that are connected by a one-way train.
\end{enumerate}
These services are semantically described using the concepts in the ontology by
the following LAV mappings:
\begin{alignat*}{1}
\nationalFlight(x,y)\ &\qrule\ \flight(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
\AAflight(x,y)\       &\qrule\ \flight(x,y,\AA)\,, \\
\UAflight(x,y)\       &\qrule\ \flight(x,y,\UA)\,, \\
\onestop(x,z)\        &\qrule\ \flight(x,y,t),\,\flight(y,z,t)\,, \\
\toPA(x)\             &\qrule\ \flight(x,\PA,\AA)\,, \\
%\onestopPA(x,y)\      &\qrule\ \flight(x,y),\,\flight(y,\PA)\,, \\
\fromLA(x)\           &\qrule\ \flight(\LA,x,\UA)\,, \\
\nationalTrain(x,y)\  &\qrule\ \train(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
\ATtrain(x,y)\        &\qrule\ \train(x,y,\AT)\,, \\
\UPtrain(x,y)\        &\qrule\ \train(x,y,\UP)\,.
\end{alignat*}
Observe that each tuple produced by each service satisfies the semantic description
expressed by the body of the rule. For example, the tuples that satisfy the predicate
$\nationalFlight(x,y)$ also meet the conjunctive formula:
\[ \exists t(\flight(x,y,t)\ \land\ \UScity(x)\ \land\ \UScity(y))\,. \]
However, there may be tuples that satisfy this formula that are not
produced by the service represented by $\nationalFlight(x,y)$.
These elements make up the integration framework $IS$. In the rest of this section.
we give examples of possible user requests and their solutions.

Suppose that a user is interested in identifying the services able to retrieve
one-stop round trips from a US city $x$ to any city $y$ in the world. Notice
that the trip from $x$ to $y$ stops at a city $u$, that the back trip from $y$
to $v$ stops at a city $v$, and that $u$ may not be equal to $v$.
This request can be modeled as the following connjunctive query:
\[ Q(x,u,y,v)\ \qrule\ \UScity(x),\,\trip(x,u),\,\trip(u,y),\,\trip(y,v),\,\trip(v,x)\,. \]
Any rewriting of the ontology predicates in terms of the services correspond
to a \emph{composition of services} that implements the user request.
For example, the following rewriting is a valid solution to the request:
\begin{alignat*}{1}
I(x,u,y,v)\ \qrule\ &\nationalFlight(x,u),\,\toPA(u),\, \\
                    &\onewayFlight(\PA,v),\,\nationalFlight(v,x)\,. 
\end{alignat*}
But, the following two rewritings are not valid solutions:
\begin{alignat*}{1}
 I'(x,u,y,v)\  &\qrule\ \nationalFlight(x,u),\,\toPA(u),\,\fromLA(v),\,\nationalFlight(v,x)\,, \\
I''(x,u,y,v)\  &\qrule\ \onestop(x,y),\,\onewayFlight(y,v),\,\nationalFlight(v,x)\,.
\end{alignat*}
The first is not valid because it maps the query variable $y$ into two different
constants \PA\ and \LA\ that denote different cities, and the second rewriting is
not valid because the service $\onestop(x,y)$ does not receive as input, or produce
as output, the middle city $u$ where the flight from $x$ to $y$ must stop.

@@

To illustrate the use of preferences, consider the following request:
\[ Q(x,y)\ \qrule\ \trip(\LA,x),\,\trip(x,\NY),\,\trip(\NY,y),\,\trip(y,\LA) \]
that looks for round-trips between Los Angeles and New York such that
each direction is a one-stop trip. Observe that the query is posed in 
a way that there are no restrictions whatsoever on the use of planes or
trains for any leg of the trip. However, as is typical, users have 
preferences about using planes or trains.
For this example, we study four different scenarios for user's preferences
and show how to model them in the proposed framework:
\begin{enumerate}[P1.]
\item The user prefers to flight rather than to travel by train. This can be
      modeled by assigning a high reward to the symbol $\flight$. Likewise, a
      preference of trains over airplanes can be modeled by assigning a high
      reward to the symbol $\train$.
\item The user is indiferent with respect to trains or airplanes, yet
      s/he does not want to mix both. This preference is an at-most-one preference
      over the set $\{\flight,\train\}$ that corresponds to the formula
      $\neg\flight \lor \neg\train$ and a cost for the violation of the
      preference.
\item If the user travels by airplane, s/he prefers to always use the same
      airline (independently of the airline). This preference can be modeled
      with the formula $\neg\AAflight \lor \neg\UAflight$ together with a
      cost. Additionally, the other means of air transportation should be
      `disabled' since they may return flights operated by any airline;
      e.g., add the constraint $\neg\nationalFlight$ with a high cost.
\item Finally, if the user travels by airplane, s/he prefers to use $\UA$.
      This is a complex preference that can be modeled with the formula
\[ (\flight \Rightarrow \UAflight) \land (\neg\UAflight \lor \neg\AAflight)\,. \]
      The first part says that if a leg of the trip is done by plane, then $\UA$
      must be used, while the second part says that whenever $\UA$ is used,
      $\AA$ should not be used. Also, the services that do not guarantee
      airline operators should be disabled as in the previous case.
\end{enumerate}
All these preferences correspond to formulas over the propositional language
$\L(IS)$. The formulas for the last three preferences can be treated as
hard constraints if the are associated with infinite cost, or as soft
constraints meaning that the user prefers (but is not limited to) solutions
that do not violate the preferences.

@@

We incorporate the subsumtion relation into \mcdsat\ by means
of the dependency graph $G(D)$. Once the graph is built using
the subsumption rules, its transitive closure is computed along
with the bindings associated with each edge: edges generated by
the transitive closure have labels that correspond to the union
of the bindings along the edges that generate this edge (if the
set of bindings is inconsistent, then the label is assigned the
binding \textbf{false}).
These labels are unique and well defined as $G(D)$ is assumed
to be a forest of trees. Once the transitive closure $G(D)^*$
is computed, all edges with inconsistent labels can be dropped.
The transitive closure is then used to extend the rules in the
logical theory that permit the cover of relational symbols in
the query with symbols in the views: a predicate $P$ is allowed
to cover a predicate $R$ whenever there is an edge from $P$ to
$R$ in $G(D)^*$, yet if this covering becomes active, the 
bindings associated with it become active as well.

@@

In order to account for the preferences, we shall use the
concepts literal-ranking functions and best-ranked models for
propositonal logic. A literal ranking function $r$ is a function
that assign ranks (real numbers) to propositional literals.
Given a literal-ranking function $r$, the rank $r(\omega)$ of
a model $\omega$ is the aggregation of the ranks for each literal
made true by the model: $r(\omega)=\sum_{\omega\vDash\ell} r(\ell)$.
Thus, the models can be ordered by their rank and the best-ranked
models are the models with minimum rank.
Some model enumerators like c2d can be used to compute all the
best-ranked models of a propositional theory.
Likewise, Weighted-Max-SAT solvers such as MiniMaxSAT~\cite{HerasetalJAIR2008}
can be used to find a best ranked model of a propositional theory.

For SSPs, we accomodate the preferences by using a suitably defined
literal-ranking function $r^*$ and by computing best-ranked models
\cite{darwiche:weighted}.
First, a new propositional variable is created for each relational
symbol in the ontology and services signatures along with clauses
that turn this proposition true whenever the corresponding symbols
become active (true). Second, for each preference $\pi=\tup{\varphi,c}$,
a new propositional symbol $p_\pi$ is created along with the
formula $p_\pi \Leftrightarrow \varphi$. Thus, $p_\pi$ is true
iff $\varphi$ is satisfied in the model (rewriting).
Finally, the literal-ranking function $r^*$ is defined as 
$r^*(\neg p_\pi)=c$ for each such preference.
Clearly, the rank of a model correspond to the sum of the costs
associated with the preferences violated by the model, and thus
a best-ranked model correspond to a rewriting of minimum regret.

@@

Finally, to study the effects of the ontology size, we considered the first benchmark and built 
ontologies where the predicates $\flight(x,y,z)$ and $\UScity(x)$ were leaves; ontologies were binary trees comprised of 2, 3, 4, 5, 6 and 7 levels.  The user request has the form: 
\[ Q(x_1,x_2,x_3,x_4)\ \qrule\  \UScity(x_1),\,\trip(x_1,x_2),\,\trip(x_2,x_3),\,\trip(x_3,x_4),\,\trip(x_4,x_1).\]
where $trip(x,y)$ was the root of the ontologies.  The results are reported in panel 
(b) of Fig.~\ref{fig:plot4}. The time to enumerate all the rewritings suggests a sub-exponential behavior, while the time to find the best models does not grow monotonically. On the other hand, 
the size of the theories and the number of models grow exponentially.  For example, the d-DNNF for 
this user request and the ontology with 7 levels is of size 1.7Mb and has
$2.6\times 10^8$ models. ÊThe time to enumerate all models is about 147 hours.

>>>>

We propose a solution that extends the recent approach of Arvelo et
al.\ \cite{arvelo:aaai06} which is based on the efficient enumeration
of models of a propositional logic  theory. Similarly, given an
instance of SSP, a logical theory is constructed such  that each model
of the theory encodes a valid rewriting; the theory can be constructed
in polynomial time from the instance of the SSP. All valid
rewritings  are obtained by enumerating all models of the theory in
linear time by using  off-the-shelf model enumeration solvers such as
c2d~\cite{c2d} and Relsat~\cite{relsat}; off-the-shelf SAT solvers
such as Minisat~\cite{minisat} can be used to produce all the minimal
rewritings. Thus, we exploit known properties of modern SAT solvers
that  implement non-chronological backtracking via clause learning,
caching of common  subproblems and heuristics, to provide an efficient
and scalable solution to SSP.

@@

Our setting is based on the assumption that services are
semantically described in terms of a domain ontology. 
Formally, the integration framework corresponds to a tuple
$IS=\tup{D,S,M}$ where $D$ is the domain ontology, $S$ describes
the available services, and $M$ is the collection of LAV mappings
that define the services in $S$ as views on the ontology $D$. 
On the other hand, the user input $R=\tup{Q,P}$ consists of a
request for composition of services, described as a conjunctive
query $Q$ over the ontology, and a set $P$ of preferences for
the requested composition.

@@

Although limited in appearance, rules of this form are quite expressive
as they allow us to specify diverse relationships among predicates.
For example,
\begin{itemize}
\item \textbf{Hierarchy of classes and subclasses} (or types and subtypes):
classes are specified with unary predicates. A subclass relationship
among two classes is directly specified with a simple rule; e.g., the
rule $\textit{penguin}(x) \orule \textit{bird}(x)$ declares that every
penguin is a bird.
\item \textbf{Subrelations via specialization:} A subrelation of $R^r$
can be specified by constraining another relation $P^s$ ($r<s$). 
For example, the rule:
\begin{equation}
\label{eq:QE2}
\textit{descendant}(\textit{Elizabeth II},x)\ \orule\ \textit{noble}(x)
\end{equation}
tells that any descendant of Queen Elizabeth II is a noble.
\item \textbf{Indirect subsumption:} it is even possible to specify
a subrelation via another seemingly unrelated predicate. For example,
the rule:
\[ \textit{citizen-of}(x,\textit{Montreal})\ \orule\ \textit{lives-in}(x,\textit{Canada}) \]
says that when the second argument of `CitizenOf' is fixed to the
constant `Montreal', the tuples in the relation `CitizenOf' are
contained in the set of tuples in the relation `LivesIn' whose
second component is `Canada'.
\end{itemize}
However, we require that the \emph{dependency graph} for the ontology
to be a \emph{forest of trees}. The dependency graph $G(D)$ is a
labeled directed graph that is constructed as follows: the nodes
are the relational symbols in the signature $\sigma$, and there
is an edge $(R,P)$ in the graph iff there is an rule of the
form: \eqref{eq:orule} that is labeled with the \emph{bindings}
induced by the rule. For example, if $\textit{descendant}(x,y)$
and $\textit{noble}(z)$ are two predicates in the signature
and there is the rule \eqref{eq:QE2}, there edge from 
Descendant to Noble is labeled with the bindings
$\{x=\textit{Elizabeth II},y=z\}$.

@@

\Omit{
$S$ is a signature  defined as a tuple $<$$S_1^{j1}$,$\dots$,
$S_q^{jn}$, $d_1$,$\dots$,$d_k$$>$, where $S_i^{ji}$ corresponds
to a service predicate of arity $ji$ and $d_i$ is a constant.  
Mappings in $M$ correspond to semantic descriptions expressed
following the  Local as View paradigm (LAV). We assume that
services are not complete, i.e., every tuple produced by the
source in the head of the rule must satisfy the body, but not
vice versa. LAV rules are commonly used in integration systems
in order to scale up to new sources, in contrast to the Global
As View paradigm (GAV) where domain ontology predicates are
defined in terms of conjunctive formulas in the service predicates.
In addition, in LAV mappings, services are treated independently
and the integration system easily scales up to the evolution of
the available services. However, the problem of rewriting a query
on the domain ontology into a set of queries on the available
services, corresponds to the problem of rewriting queries using
views which is known to be NP-complete~\cite{RajaramanSU95,Ullman00}.  
}

@@

Let us consider a simple travel-information system that contains
information about flight and train trips between cities and information
about which cities are in the US. The domain ontology is comprised
of the predicates $\trip^2$, $\AAflight^2$, $\UAflight^2$, $\ATtrain^2$,
$\UPtrain^2$, $\flight^3$, $\train^3$ and $\UScity^1$, and the constants
$\AA$, $\UA$, $\AT$, $\UP$, $\LA$, $\NY$, $\PA$.
The first predicate relates cities $(x,y)$ if there is a direct
trip either by plane or train between them.
The predicates $\AAflight$ and $\UAflight$ relate cities that are
connected by direct flights either with American (AA) or United (UA).
The predicates $\ATtrain$ and $\UPtrain$ relate cities that are
connected by direct train trips either with Amtrak (AT) or Union
Pacific Railway (UP). The flight predicate relates $(x,y,t)$ whenever
there is a direct flight from $x$ to $y$ operated by airline $t$,
and similarly for $\train$. Finally, $\UScity$ indicates when a
given city is a US city or not.
There are various subsumption relationships among these predicates
that are captured by the following axioms:
\begin{alignat*}{1}
\flight(x,y,t)\  &\orule\ \trip(x,y)\,, \\
\train(x,y,t)\   &\orule\ \trip(x,y)\,.
\end{alignat*}
For the services, assume that the available data sources on the
Internet contain the following information:
\begin{enumerate}[--]
\item $\nationalFlight(x,y)$ relates two US cities that are connected by a direct flight,
\item $\onewayFlight(x,y)$ relates two cities that are connected by a one-way flight,
\item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
\item $\toPA(x)$ tells if there is a direct flight from $x$ to Paris,
\item $\onestopPA(x,y)$ if there is a flight from $x$ to Paris with a stop at $y$, 
\item $\fromLA(x)$ tells if there is a flight from Los Angeles to $x$,
\item $\nationalTrain(x,y)$ relates US cities that are connected by a direct train, and
\item $\onewayTrain(x,y)$ relates US cities that are connected by a one-way train.
\end{enumerate}
These services are semantically described using the concepts in the ontology by
the following LAV mappings:
\begin{alignat*}{1}
\nationalFlight(x,y)\ &\qrule\ \flight(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
\AAflight(x,y)\       &\qrule\ \flight(x,y,\AA)\,, \\
\UAflight(x,y)\       &\qrule\ \flight(x,y,\UA)\,, \\
\onestop(x,z)\        &\qrule\ \flight(x,y,t),\,\flight(y,z,t)\,, \\
\toPA(x)\             &\qrule\ \flight(x,\PA,\AA)\,, \\
%\onestopPA(x,y)\      &\qrule\ \flight(x,y),\,\flight(y,\PA)\,, \\
\fromLA(x)\           &\qrule\ \flight(\LA,x,\UA)\,, \\
\nationalTrain(x,y)\  &\qrule\ \train(x,y,t),\,\UScity(x),\,\UScity(y)\,, \\
\ATtrain(x,y)\        &\qrule\ \train(x,y,\AT)\,, \\
\UPtrain(x,y)\        &\qrule\ \train(x,y,\UP)\,.
\end{alignat*}
Observe that each tuple produced by each service satisfies the semantic description
expressed by the body of the rule. For example, the tuples that satisfy the predicate
$\nationalFlight(x,y)$ also meet the conjunctive formula:
\[ \exists t(\flight(x,y,t)\ \land\ \UScity(x)\ \land\ \UScity(y))\,. \]
However, there may be tuples that satisfy this formula that are not
produced by the service represented by $\nationalFlight(x,y)$.
These elements make up the integration framework $IS$. In the rest of this section.
We give examples of possible user requests and their solutions.

Suppose that a user is interested in identifying the services able to retrieve
one-stop round trips from a US city $x$ to any city $y$ in the world. Notice
that the trip from $x$ to $y$ stops at a city $u$, that the back trip from $y$
to $v$ stops at a city $v$, and that $u$ may not be equal to $v$.
This request can be modeled as the following conjunctive query:
\[ Q(x,u,y,v)\ \qrule\ \UScity(x),\,\trip(x,u),\,\trip(u,y),\,\trip(y,v),\,\trip(v,x)\,. \]
Any rewriting of the ontology predicates in terms of the services correspond
to a \emph{composition of services} that implements the user request.
For example, the following rewriting is a valid solution to the request:
\begin{alignat*}{1}
I(x,u,y,v)\ \qrule\ &\nationalFlight(x,u),\,\toPA(u),\, \\
                    &\onewayFlight(\PA,v),\,\nationalFlight(v,x)\,. 
\end{alignat*}
But, the following two rewritings are not valid solutions:
\begin{alignat*}{1}
 I'(x,u,y,v)\  &\qrule\ \nationalFlight(x,u),\,\toPA(u),\,\fromLA(v),\,\nationalFlight(v,x)\,, \\
I''(x,u,y,v)\  &\qrule\ \onestop(x,y),\,\onewayFlight(y,v),\,\nationalFlight(v,x)\,.
\end{alignat*}
The first is not valid because it maps the query variable $y$ into two different
constants \PA\ and \LA\ that denote different cities, and the second rewriting is
not valid because the service $\onestop(x,y)$ does not receive as input, or produce
as output, the middle city $u$ where the flight from $x$ to $y$ must stop.

@@

To illustrate the use of preferences, consider the following request:
\[ Q(x,y)\ \qrule\ \trip(\LA,x),\,\trip(x,\NY),\,\trip(\NY,y),\,\trip(y,\LA) \]
that looks for round-trips between Los Angeles and New York such that
each direction is a one-stop trip. Observe that the query is posed in 
a way that there are no restrictions whatsoever on the use of planes or
trains for any leg of the trip. However, as is typical, users have 
preferences about using planes or trains.
For this example, we study four different scenarios for user's preferences
and show how to model them in the proposed framework:
\begin{enumerate}[P1.]
\item The user prefers to flight rather than to travel by train. This can be
      modeled by assigning a high reward to the symbol $\flight$. Likewise, a
      preference of trains over airplanes can be modeled by assigning a high
      reward to the symbol $\train$.
\item The user is indifferent with respect to trains or airplanes, yet
      s/he does not want to mix both. This preference is an at-most-one preference
      over the set $\{\flight,\train\}$ that corresponds to the formula
      $\neg\flight \lor \neg\train$ and a cost for the violation of the
      preference.
\item If the user travels by airplane, s/he prefers to always use the same
      airline (independently of the airline). This preference can be modeled
      with the formula $\neg\AAflight \lor \neg\UAflight$ together with a
      cost. Additionally, the other means of air transportation should be
      `disabled' since they may return flights operated by any airline;
      e.g., add the constraint $\neg\nationalFlight$ with a high cost.
\item Finally, if the user travels by airplane, s/he prefers to use $\UA$.
      This is a complex preference that can be modeled with the formula
\[ (\flight \Rightarrow \UAflight) \land (\neg\UAflight \lor \neg\AAflight)\,. \]
      The first part says that if a leg of the trip is done by plane, then $\UA$
      must be used, while the second part says that whenever $\UA$ is used,
      $\AA$ should not be used. Also, the services that do not guarantee
      airline operators should be disabled as in the previous case.
\end{enumerate}
All these preferences correspond to formulas over the propositional language
$\L(IS)$. The formulas for the last three preferences can be treated as
hard constraints if the are associated with infinite cost, or as soft
constraints meaning that the user prefers (but is not limited to) solutions
that do not violate the preferences.

@@

We incorporate the subsumption relation into \mcdsat\ by means
of the dependency graph $G(D)$. Once the graph is built using
the subsumption rules, its transitive closure is computed along
with the bindings associated with each edge: edges generated by
the transitive closure have labels that correspond to the union
of the bindings along the edges that generate this edge (if the
set of bindings is inconsistent, then the label is assigned the
binding \textbf{false}).
These labels are unique and well defined as $G(D)$ is assumed
to be a forest of trees. Once the transitive closure $G(D)^*$
is computed, all edges with inconsistent labels can be dropped.
The transitive closure is then used to extend the rules in the
logical theory that permit the cover of relational symbols in
the query with symbols in the views: a predicate $P$ is allowed
to cover a predicate $R$ whenever there is an edge from $P$ to
$R$ in $G(D)^*$, yet if this covering becomes active, the 
bindings associated with it become active as well.

@@

In order to account for the preferences, we shall use the
concepts literal-ranking functions and best-ranked models for
propositional logic. A literal ranking function $r$ is a function
that assign ranks (real numbers) to propositional literals.
Given a literal-ranking function $r$, the rank $r(\omega)$ of
a model $\omega$ is the aggregation of the ranks for each literal
made true by the model: $r(\omega)=\sum_{\omega\vDash\ell} r(\ell)$.
Thus, the models can be ordered by their rank and the best-ranked
models are the models with minimum rank.
Some model enumerators like c2d can be used to compute all the
best-ranked models of a propositional theory.
Likewise, Weighted-Max-SAT solvers such as MiniMaxSAT~\cite{HerasetalJAIR2008}
can be used to find a best ranked model of a propositional theory.

For SSPs, we accommodate the preferences by using a suitably defined
literal-ranking function $r^*$ and by computing best-ranked models
\cite{darwiche:weighted}.
First, a new propositional variable is created for each relational
symbol in the ontology and services signatures along with clauses
that turn this proposition true whenever the corresponding symbols
become active (true). Second, for each preference $\pi=\tup{\varphi,c}$,
a new propositional symbol $p_\pi$ is created along with the
formula $p_\pi \Leftrightarrow \varphi$. Thus, $p_\pi$ is true
iff $\varphi$ is satisfied in the model (rewriting).
Finally, the literal-ranking function $r^*$ is defined as 
$r^*(\neg p_\pi)=c$ for each such preference.
Clearly, the rank of a model correspond to the sum of the costs
associated with the preferences violated by the model, and thus
a best-ranked model correspond to a rewriting of minimum regret.

@@

Finally, to study the effects of the ontology size, we considered the first benchmark and built 
ontologies where the predicates $\flight(x,y,z)$ and $\UScity(x)$ were leaves; ontologies were binary trees comprised of 2, 3, 4, 5, 6 and 7 levels.  The user request has the form: 
\[ Q(x_1,x_2,x_3,x_4)\ \qrule\  \UScity(x_1),\,\trip(x_1,x_2),\,\trip(x_2,x_3),\,\trip(x_3,x_4),\,\trip(x_4,x_1).\]
where $trip(x,y)$ was the root of the ontologies.  The results are reported in panel 
(b) of Fig.~\ref{fig:plot4}. The time to enumerate all the rewritings suggests a sub-exponential behavior, while the time to find the best models does not grow monotonically. On the other hand, 
the size of the theories and the number of models grow exponentially.  For example, the d-DNNF for 
this user request and the ontology with 7 levels is of size 1.7Mb and has
$2.6\times 10^8$ models. The time to enumerate all models is about 147 hours.
